---
title: "p8105_hw3_sc5351"
author: "Shaoyu Chen"
date: "2023-10-13"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggridges)
library(ggplot2)
library(knitr)
library(patchwork)
```

##Problem 1
```{r include=FALSE}
library(p8105.datasets)
knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

```{r}
data("instacart")
instacart = 
  instacart |> 
  as_tibble()
```

The size of the dataset is `r nrow(instacart)`rows and  `r ncol(instacart)` columns. The important varables are `order_id`, `product_id`, `add_to_cart_order`, `reordered` and `user_id`.There are several order-level variables, describing the day and time of the order, and number of days since prior order. Then there are several item-specific variables, describing the product name (e.g. Yogurt, Avocado), department (e.g. dairy and eggs, produce), and aisle (e.g. yogurt, fresh fruits), and whether the item has been ordered by this user in the past. In total, there are `r instacart |> select(product_id) |> distinct() |> count()` products found in `r instacart |> select(user_id, order_id) |> distinct() |> count()` orders from `r instacart |> select(user_id) |> distinct() |> count()` distinct users.

Below is a table summarizing the number of items ordered from aisle. In total, there are 134 aisles, with fresh vegetables and fresh fruits holding the most items ordered by far.

```{r}
instacart |> 
  count(aisle) |> 
  arrange(desc(n))
```

Below is a plot that shows the number of items ordered in each aisle.
```{r}
instacart |> 
  count(aisle) |> 
  filter(n > 10000) |> 
  mutate(aisle = fct_reorder(aisle, n)) |> 
  ggplot(aes(x = aisle, y = n)) + 
  geom_point() + 
  labs(title = "Number of items ordered in each aisle") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
```

Below is a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. 
```{r}
instacart |> 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) |>
  group_by(aisle) |> 
  count(product_name) |> 
  mutate(rank = min_rank(desc(n))) |> 
  filter(rank < 4) |> 
  arrange(desc(n)) |>
  knitr::kable()
```

Below is a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week.
```{r}
instacart |>
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) |>
  group_by(product_name, order_dow) |>
  summarize(mean_hour = mean(order_hour_of_day)) |>
  pivot_wider(
    names_from = order_dow, 
    values_from = mean_hour) |>
  knitr::kable(digits = 2)
```

##Proble 2
Data cleaning
```{r}
data("brfss_smart2010")
brfss_data <-
  brfss_smart2010|>
  janitor::clean_names()|>
  rename("state"="locationabbr" , "location"="locationdesc") |>
  filter(topic=='Overall Health')|>
  filter(response %in% c("Excellent","Very good","Good","Fair","Poor"))|>
  mutate(response = factor(response, levels = c("Poor", "Fair", "Good", "Very good", "Excellent")))
brfss_data
```
States were observed at 7 or more locations in 2002.
```{r}
state_2002 =
  brfss_data|>
  filter(year == "2002")|>
  group_by(year,state)|>
  summarize(count_locations = n_distinct(location))|>
  filter(count_locations > 6)|>
  knitr::kable(digit = 2)
state_2002
```

States were observed at 7 or more locations in 2010.
```{r}
state_2010 = 
  brfss_data|>
  filter(year == "2010")|>
  group_by(year,state)|>
  summarize(count_locations = n_distinct(location))|>
  filter(count_locations > 6)|>
  knitr::kable(digit = 2)
state_2010
```
In 2002, CT, FL, MA, NC, NJ, and PA (totally 6) were observed at 7 or more locations.
In 2010, FL, MA, MD, NC, NE, NJ, NY, OH, PA, SC, TX and WA (totally 14) were observed at 7 or more locations.

Then construct a dataset that is limited to Excellent responses, and contains, year, state, and a variable that averages the data_value across locations within a state. 
```{r}
spaghetti_data <- brfss_data |> 
  filter(response == "Excellent") |>
  group_by(year, state) |>
  summarise(average_value = mean(data_value, na.rm = TRUE), .groups = "drop")
spaghetti_data|>
  ggplot(aes(x = year, y = average_value, group = state)) + 
  geom_line(aes(color = state), alpha = .5) + 
  labs(
    title = "Average Value of Excellent Responses",
    x = "Year",
    y = "Average Value",
    caption = "Data from brfss_data"
  ) +
  theme(legend.position = "right")
```

Finally, make a two-panel plot showing, for the years 2006, and 2010, distribution of data_value for responses (“Poor” to “Excellent”) among locations in NY State.
```{r}
plot_data <- brfss_data |>
  filter(year %in% c("2006", "2010"),state == "NY")|>
  select(year, state, location, response, data_value)
plot_data |>
  ggplot(aes(x = response,y = data_value, color=response))+
  geom_boxplot()+
  labs(
    x = "Response",
    y = "Data Value",
    title = "Distribution of data_value for Responses in NY State in 2006 and 2010"
  ) +
  facet_grid(.~year) +
  theme(legend.position = "right")
```

##Problem 3
Load, tidy, merge, and otherwise organize the data sets.
```{r}
covar_data =
  read_csv("./data/nhanes_covar.csv", skip = 4) |>
  janitor::clean_names() |>
  filter(age >= 21) |>
  drop_na()|>
  mutate(
    sex = case_when(
    sex == 1 ~ "Male",
    sex == 2 ~ "Female"
  ),
  education = case_when(
    education == 1 ~ "Less than high school",
    education == 2 ~ "High school equivalent",
    education == 3 ~ "More than high school"
  )) 
covar_data
```
```{r}
accel_data <-
  read_csv("./data/nhanes_accel.csv")|>
  janitor::clean_names()
accel_data
```
Produce a reader-friendly table for the number of men and women in each education category, and create a visualization of the age distributions for men and women in each education category.


